Simple and robust inference of connectomes using partial correlation coefficients

### 2014-06-11 article.tex
wording + f_3 -> tau

Connectivity information between two neurons may be retrieved from successive
and slightly delayed burst events. Direct links
between two neurons should have shorter delay between fluorescence
peaks than undirected links. However, the fluorescence signal has a slow
decay compare to its rise. Therefore, we filter those unrelevant information
with a hard thresholding filter
\begin{align}
X^{\prime\prime\prime}_{t,i} &=
{X^{\prime\prime}_{t,i}}^{0.9} I(X^{\prime\prime}_{t,i} \geq f_3) \forall i
\label{eq:hard-treshold-filter}
\end{align}
where $f_3 \geq 0$ is a parameter of the hard-threshold filter and the $0.9$
exponent reduce the difference between peaks values.

@@

% Speak about why averaging
The quality of the signal filtering and network inference is dependent upon
good hyper-parameter selection, which are themselves dependent on the properties
of the targeted network and measurement conditions, e.g. signal to noise ratio.
Ensemble methods are known to improve robustness with respect to parameter
selection [todo cite] and improve performance [todo cite].
A weighted average $y_{stack}$ of partial correlation matrices was made
by varying uniformly the hard-threshold value (parameter $f_3$) and the design
of the low-pass filter (coefficients $c_l$). The stacked connectivity score
matrix $y_{stacking}$ is given by
\begin{align}
y_{stack} =  \dfrac{1}{\sum_{k\in\{1,..,4\}} w_k |T_k|} \sum_{k=1}^{k=4}  w_k \sum_{q \in T_k} G^{q,k}(X_t^i)
\end{align}
where $f(X, k, f_3)$ gives the predicted adjacency matrix, i.e. applies the
filtering and inference methodologies (as described
in sections \ref{sec:filter} and \ref{sec:inference}), with a  given value
$f_3$ for the hard-threshold among a set of thresholds $T_k$ of size $|T_k|$
and the $k^{th}$ low-pass filter pattern associated to a weight $w_k$ whose
values are $w_1 = 0.01 , w_2 =1, w_3 = 0.7, w_4 = 0.9$.

@@

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link (see a more detailed discussion about the
chosen metric in section \ref{sec:metric}). Therefore, performance might be
improved by predicting an asymmetric adjacency matrix. Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $z_{ij}$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
timesteps:
\[
z_{ij} = \sum_{t=1}^n \mathbb{1}(X_{t+1,j} \ge X_{t, i}) \cdot
\mathbb{1}(f_1\le X_t^i \le f_2) -  \mathbb{1}(X_{t+1,i} \ge X_{t, j}) \cdot
\mathbb{1}(f_1 \le X_t^j \le f_2)
\]
where $f_1$ and $f_2$ are parameters of the method and $I$ is the indicator
function. \footnote{Only the filter defined by Equations \ref{eq:low-pass},
\ref{eq:high-pass} and \ref{eq:hard-treshold-filter}}.

>>>>

Connectivity information between two neurons may be retrieved from successive
and slightly delayed burst events. Direct links
between two neurons should have shorter delay between fluorescence
peaks than undirected links. However, the fluorescence signal has a slow
decay compare to its rise. Therefore, we filter those unrelevant information
with a hard thresholding filter
\begin{align}
X^{\prime\prime\prime}_{t,i} &=
{X^{\prime\prime}_{t,i}}^{0.9} I(X^{\prime\prime}_{t,i} \geq \tau) \forall i
\label{eq:hard-treshold-filter}
\end{align}
where $\tau \geq 0$ is a parameter of the hard-threshold filter and the $0.9$
exponent reduce the difference between peaks values.

@@

% Speak about why averaging
The quality of the signal filtering and network inference is dependent upon
good hyper-parameter selection, which are themselves dependent on the properties
of the targeted network and measurement conditions, e.g. signal to noise ratio.
Ensemble methods are known to improve robustness with respect to parameter
selection [todo cite] and improve performance [todo cite].
A weighted average $y_{stack}$ of partial correlation matrices was made
by varying uniformly the hard-threshold value (parameter $\tau$) and the design
of the low-pass filter (coefficients $c_l$). The stacked connectivity score
matrix $y_{stacking}$ is given by
\begin{align}
y_{stack} =  \dfrac{1}{\sum_{k\in\{1,..,4\}} w_k |T_k|} \sum_{k=1}^{k=4}  w_k \sum_{q \in T_k} G^{q,k}(X_t^i)
\end{align}
where $f(X, k, \tau)$ gives the predicted adjacency matrix, i.e. applies the
filtering and inference methodologies (as described
in sections \ref{sec:filter} and \ref{sec:inference}), with a  given value
$\tau$ for the hard-threshold among a set of thresholds $T_k$ of size $|T_k|$
and the $k^{th}$ low-pass filter pattern associated to a weight $w_k$ whose
values are $w_1 = 0.01 , w_2 =1, w_3 = 0.7, w_4 = 0.9$.

@@

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link (see a more detailed discussion about the
chosen metric in section \ref{sec:metric}). Therefore, performance might be
improved by predicting an asymmetric adjacency matrix. Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $z_{ij}$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
timesteps:
\[
z_{ij} = \sum_{t=1}^n \mathbb{1}(X_{t+1,j} \ge X_{t, i}) \cdot
\mathbb{1}(f_1\le X_t^i \le f_2) -  \mathbb{1}(X_{t+1,i} \ge X_{t, j}) \cdot
\mathbb{1}(f_1 \le X_t^j \le f_2)
\]
where $f_1$ and $f_2$ are parameters of the method and $I$ is the indicator
function. Prior computing $z_{ij}$ only the filter defined by Equations
\ref{eq:low-pass}, \ref{eq:high-pass} and \ref{eq:hard-treshold-filter} were
used.
